---
title: "HW1 Peer Assessment"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part A. ANOVA

Additional Material: ANOVA tutorial

https://datascienceplus.com/one-way-anova-in-r/

Jet lag is a common problem for people traveling across multiple time zones, but people can gradually adjust to the new time zone since the exposure of the shifted light schedule to their eyes can resets the internal circadian rhythm in a process called “phase shift”. Campbell and Murphy (1998) in a highly controversial study reported that the human circadian clock can also be reset by only exposing the back of the knee to light, with some hailing this as a major discovery and others challenging aspects of the experimental design. The table below is taken from a later experiment by Wright and Czeisler (2002) that re-examined the phenomenon. The new experiment measured circadian rhythm through the daily cycle of melatonin production in 22 subjects randomly assigned to one of three light treatments. Subjects were woken from sleep and for three hours were exposed to bright lights applied to the eyes only, to the knees only or to neither (control group). The effects of treatment to the circadian rhythm were measured two days later by the magnitude of phase shift (measured in hours) in each subject’s daily cycle of melatonin production. A negative measurement indicates a delay in melatonin production, a predicted effect of light treatment, while a positive number indicates an advance.

Raw data of phase shift, in hours, for the circadian rhythm experiment

|Treatment|Phase Shift (hr)                            |
|:--------|:-------------------------------------------|
|Control  |0.53, 0.36, 0.20, -0.37, -0.60, -0.64, -0.68, -1.27|
|Knees    |0.73, 0.31, 0.03, -0.29, -0.56, -0.96, -1.61       |
|Eyes     |-0.78, -0.86, -1.35, -1.48, -1.52, -2.04, -2.83    |

## Question A1 - 3 pts

Consider the following incomplete R output:

|Source|Df |Sum of Squares|Mean Squares|F-statistics|p-value|
|:----:|:-:|:------------:|:----------:|:----------:|:-----:|
|Treatments|2|7.224|3.6122|7.289|0.004|
|Error|19|9.415|0.496| | |
|TOTAL|21|16.639| | | |

Fill in the missing values in the analysis of the variance table.Note: Missing values can be calculated using the corresponding formulas provided in the lectures, or you can build the data frame in R and generate the ANOVA table using the aov() function. Either approach will be accepted.

Answer:
```{r}
control_group <- c(0.53, 0.36, 0.20, -0.37, -0.60, -0.64, -0.68, -1.27)
control_label <- rep("control",length(control_group))
control <- cbind(control_label,control_group)
knees_group <- c(0.73, 0.31, 0.03, -0.29, -0.56, -0.96, -1.61)
knees_label <- rep("knees",length(knees_group))
knees <- cbind(knees_label,knees_group)
eyes_group <- c(-0.78, -0.86, -1.35, -1.48, -1.52, -2.04, -2.83)
eyes_label <- rep("eyes",length(eyes_group))
eyes <- cbind(eyes_label,eyes_group)
df <- as.data.frame(rbind(control, knees, eyes))
colnames(df) <- c('treatment', 'phase_shift')
model <- aov(phase_shift ~ treatment, data = df)
summary(model)

```


## Question A2 - 3 pts

Use $\mu_1$, $\mu_2$, and $\mu_3$  as notation for the three mean parameters and define these parameters clearly based on the context of the topic above  (i.e. explain what  $\mu_1$, $\mu_2$, and $\mu_3$ mean in words in the context of this problem). Find the estimates of these parameters.

Answer: 
```{r}
u1 <- mean(control_group)
u2 <- mean(knees_group)
u3 <- mean(eyes_group)
u1
## -0.30875
u2
## -0.3357143
u3
## -1.551429
```
u1 is the average shift, in hours, for the control group - no bright lights exposed to either knees or eyes. The mean is -0.30875.
u2 is the average shift, in hours, for the group that only had bright lights exposed to the knees. The mean is -0.3357143.
u3 is the average shift, in hours, for the group that only had bright lights exposed to the eyes. The mean is -1.551429.

## Question A3 - 5 pts

Use the ANOVA table in Question A1 to answer the following questions:

a. **1 pts** Write the null hypothesis of the ANOVA $F$-test, $H_0$
  1. There is no difference between the means of each group. 

b. **1 pts** Write the alternative hypothesis of the ANOVA $F$-test, $H_A$
  1. There is a difference between at least one of the means of each group.

c. **1 pts** Fill in the blanks for the degrees of freedom of the ANOVA $F$-test statistic:   $F$(____, _____)
  1. (2,19)

d. **1 pts** What is the p-value of the ANOVA $F$-test?
  1. 0.00447

e. **1 pts** According the the results of the ANOVA $F$-test, does light treatment affect phase shift?  Use an $\alpha$-level of 0.05.
  1. Yes, the p-value of the F-test is less than our alpha level of 0.05. Due to that we can reject the null hypothesis and say that there is a difference between at least one of the means of each group.

# Part B. Simple Linear Regression

We are going to use regression analysis to estimate the performance of CPUs based on the maximum number of channels in the CPU.  This data set comes from the UCI Machine Learning Repository.

The data file includes the following columns:

* *vendor*: vendor of the CPU
* *chmax*: maximum channels in the CPU
* *performance*: published relative performance of the CPU

The data is in the file "machine.csv". To read the data in `R`, save the file in your working directory (make sure you have changed the directory if different from the R working directory) and read the data using the `R` function `read.csv()`.

```{r}
# Read in the data
data = read.csv("machine.csv", head = TRUE, sep = ",")
# Show the first few rows of data
head(data, 3)
```

## Question B1: Exploratory Data Analysis - 9 pts

a. **3 pts** Use a scatter plot to describe the relationship between CPU performance and the maximum number of channels. Describe the general trend (direction and form). Include plots and R-code used.

Answer: 
From the plot below, we can see, generally, as chmax increases, so does performance.
```{r}
plot(data$chmax, data$performance, ylab='performance', xlab='chmax')

```

b. **3 pts** What is the value of the correlation coefficient between _performance_ and _chmax_? Please interpret the strength of the correlation based on the correlation coefficient.

Answer:
Values of the correlation coefficient is 0.6052093. It is a positive correlation but it is medium strength correlation at best.
```{r}
cor(data$chmax, data$performance)

```

c. **2 pts** Based on this exploratory analysis, would you recommend a simple linear regression model for the relationship?
    1. Yes, looking at the scatter plot and the correlation coefficient, I'd be interested in digging more into this.

d. **1 pts** Based on the analysis above, would you pursue a transformation of the data? *Do not transform the data.*
    1. Yes, part of digging more into this would be taking the log of the data. I'd like to check distribution and understand more of how the data looks.

## Question B2: Fitting the Simple Linear Regression Model - 11 pts

Fit a linear regression model, named *model1*, to evaluate the relationship between performance and the maximum number of channels. *Do not transform the data.* The function you should use in R is:

```{r}
# Your code here...
model1 = lm(performance ~ chmax, data)
summary((model1))

```

a. **3 pts** What are the model parameters and what are their estimates?  
    1. Intercept is: 37.2252, Slope is: 3.7441

b. **2 pts** Write down the estimated simple linear regression equation.
    1. Yi = 37.2252 + 3.7441Xi + e

c. **2 pts** Interpret the estimated value of the $\beta_1$ parameter in the context of the problem.
    1. Every time the maximum channels in the CPU goes up a unit of measure, the performance of the CPU goes up 3.7441 units on average.

d. **2 pts** Find a 95% confidence interval for the $\beta_1$ parameter. Is $\beta_1$ statistically significant at this level?
    1. Yes because the p-value of 2.2e-16 is less than 0.05. It falls with in the confidence interval of 3.069251 to 4.418926.
    ```{r}
confint(model1, level=0.95)

```

e. **2 pts** Is $\beta_1$ statistically significantly positive at an $\alpha$-level of 0.01?  What is the approximate p-value of this test?
    1. Yes, the p-value is less than 0.01. The p-value is 2.2e-16.

## Question B3: Checking the Assumptions of the Model - 8 pts

Create and interpret the following graphs with respect to the assumptions of the linear regression model. In other words, comment on whether there are any apparent departures from the assumptions of the linear regression model. Make sure that you state the model assumptions and assess each one.  Each graph may be used to assess one or more model assumptions.

a. **2 pts** Scatterplot of the data with *chmax* on the x-axis and *performance* on the y-axis

```{r}
plot(data$chmax, data$performance, ylab='performance', xlab='chmax')

```

**Model Assumption(s) it checks:** Linear relationship. Correlation between variables can be seen. 

**Interpretation:** It looks that performance increases as chmax increases, but the correlation is not very strong. Data is heavily grouped together at the bottom left of the graph. Potentially using a log scale would help to identify further.

b. **3 pts** Residual plot - a plot of the residuals, $\hat\epsilon_i$, versus the fitted values, $\hat{y}_i$

```{r}
residual_plt <- resid(model1)
plot(fitted(model1), residual_plt) + abline(h=0, col="blue")

```

**Model Assumption(s) it checks:** Homoscedasticity/Constant Variance Assumption.

**Interpretation:** The points are not equal as we look a differing levels of the fitted model. Points around fitted model 500 do not have similar plotted points around fitted model 100. Shows that we have outliers.

c. **3 pts** Histogram and q-q plot of the residuals

```{r}
qqnorm(residual_plt)  ##+ qqline(residual_plt)


```
```{r}
hist(residual_plt)
```

**Model Assumption(s) it checks:** Normal distribution.

**Interpretation:** At first glance, it looks to be normally distributed based on the graphs. Although, the tails of the qq plot do curve.

## Question B4: Improving the Fit - 10 pts

a. **2 pts** Use a Box-Cox transformation (`boxCox()`) in `car()` package or (`boxcox()`) in `MASS()` package to find the optimal $\lambda$ value rounded to the nearest half integer.  What transformation of the response, if any, does it suggest to perform?
  Answer: Optimal lambda value is 0. It suggests a log transformation.
```{r}
#install.packages("car")
library(car)

b_c = boxCox(model1)
```

b. **2 pts** Create a linear regression model, named *model2*, that uses the log transformed *performance* as the response, and the log transformed *chmax* as the predictor. Note: The variable *chmax* has a couple of zero values which will cause problems when taking the natural log. Please add one to the predictor before taking the natural log of it

```{r}
new_data = data
new_data$chmax[new_data$chmax == 0] <- 1 ##satisfying the issue of zero values causing problems by turning them into 1s. 
model2 = lm(log(new_data$performance) ~ log(new_data$chmax), new_data)
summary(model2)
```

c. **2 pts** Compare the R-squared values of *model1* and *model2*.  Did the transformation improve the explanatory power of the model?
    1. Yes, the transformation did improve the explanatory power of the model. R-squared went from 0.3663 to 0.418. Adjusted R-squared also went up from 0.3632 to 0.4152.
    ```{r}
    summary(model1)
    ```
d. **4 pts** Similar to Question B3, assess and interpret all model assumptions of *model2*.  A model is considered a good fit if all assumptions hold. Based on your interpretation of the model assumptions, is *model2* a good fit?

```{r}
plot(log(data$chmax), log(data$performance), ylab='performance', xlab='chmax')
#looks moore evenly distributed than model 1

```

```{r}
 m2_residuals <- resid(model2)
plot(fitted(model2), m2_residuals) + abline(h=0)
 ## better spread than model 1 
```
```{r}
qqnorm(m2_residuals) #+ qqline(m2_residuals)
#tails are not curving as much as the did in model 1
```
```{r}
hist(m2_residuals)
# histogram looks much more normally distributed
```
Answer: 
Model 2 is a good fit, all assumptions hold.

## Question B5: Prediction - 3 pts

Suppose we are interested in predicting CPU performance when `chmax = 128`.  Please make a prediction using both *model1* and *model2* and provide the 95% prediction interval of each prediction on the original scale of the response, *performance*. What observations can you make about the result in the context of the problem?

```{r}
b5_data <- data.frame(chmax = c(128))
predict(model1, b5_data, interval = 'prediction')

```
```{r}
b5_data$chmax <- log(b5_data$chmax)
exp(predict(model2, newdata = b5_data, interval = 'prediction'))
## cannot get it to show just my 1 result
```

# Part C. ANOVA - 8 pts

We are going to continue using the CPU data set to analyse various vendors in the data set.  There are over 20 vendors in the data set.  To simplify the task, we are going to limit our analysis to three vendors, specifically, honeywell, hp, and nas.  The code to filter for those vendors is provided below.

```{r}
# Filter for honeywell, hp, and nas
data2 = data[data$vendor %in% c("honeywell", "hp", "nas"), ]
data2$vendor = factor(data2$vendor)
```

1. **2 pts** Using `data2`, create a boxplot of *performance* and *vendor*, with *performance* on the vertical axis.  Interpret the plots.  
  Answer: We can see honeywell and hp have similar means while nas is seemingly statistically significantly different than both - based on the plots visuals.Honeywell also has some outliers.
```{r}
boxplot(data=data2, data2$performance ~ data2$vendor, xlab = 'vendor', ylab = 'performance')

```

2. **3 pts** Perform an ANOVA F-test on the means of the three vendors.  Using an $\alpha$-level of 0.05, can we reject the null hypothesis that the means of the three vendors are equal?  Please interpret.
  Answer: Looking at the p-value, we can reject the null hypothesis.The F-critical value of 3.259446 is also less than the F-test value, 6.027, provided by the model.

```{r}
d2_model = aov(data2$performance ~ data2$vendor)
alpha <- 0.05
an_model <- anova(d2_model)
qf(1-alpha, an_model[1,1], an_model[2,1])
```
```{r}
summary(d2_model)
```

3. **3 pts** Perform a Tukey pairwise comparison between the three vendors. Using an $\alpha$-level of 0.05, which means are statistically significantly different from each other?
  Answer: nas-honeywell & nas-hp are statistically significantly different from each other. We can tell looking at the p-values. hp-honeywell have an extremely high p-value, so we know those are not statistically different.
```{r}
TukeyHSD(d2_model)

```
