---
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Peer Grader Guidance
Please review the student expectations for peer review grading and peer review comments.  Overall, we ask that you score with accuracy. When grading your peers, you will not only learn how to improve your future homework submissions but you will also gain deeper understanding of the concepts in the assignments. When assigning scores, consider the responses to the questions given your understanding of the problem and using the solutions as a guide. Moreover, please give partial credit for a concerted effort, but also be thorough. **Add comments to your review, particularly when deducting points, to explain why the student missed the points.** Ensure your comments are specific to questions and the student responses in the assignment.


# Background

You have been contracted as a healthcare consulting company to understand the factors on which the pricing of health insurance depends. 

## Data Description

The data consists of a data frame with 1338 observations on the following 7 variables:

1. price: Response variable ($)
2. age: Quantitative variable
3. sex: Qualitative variable
4. bmi: Quantitative variable
5. children: Quantitative variable
6. smoker: Qualitative variable
7. region: Qualitative variable


## Instructions on reading the data

To read the data in `R`, save the file in your working directory (make sure you have changed the directory if different from the R working directory) and read the data using the `R` function `read.csv()`

```{r}
insurance = read.csv("insurance.csv", head = TRUE)
```


# Question 1: Exploratory Data Analysis [15 points]

a. **4 pts** Create scatterplots of the response, *price*, against three quantitative predictors *age*, *bmi*, and *children*. Describe the general trend (direction and form) of each plot. It should be 3 separate scatter plots.
```{r}
par(mfrow=c(2,2))
plot(price~age,data=insurance,main="Price vs. Age",col="grey",pch =16)+abline(lm(price~age,data=insurance),col="red")
```
Answer: For the plot above, there looks to be a positive linear relationship. It is not a strong one.
```{r}
plot(price~bmi,data=insurance,main="Price vs. BMI",col="grey",pch =16)+abline(lm(price~bmi,data=insurance),col="red")
```
Again, there is a weak, positive, linear relationship between price and bmi. 
```{r}
plot(price~children,data=insurance,main="Price vs. Children",col="grey",pch =16)+abline(lm(price~children,data=insurance),col="red")
```
For the plot of price compared to children, there is a slight relationship but it is very flat. It is weak than the realtionship between age and bmi when compared against price.

b. **4 pts** What is the value of the correlation coefficient for each of the above pair of response and predictor variables? What does it tell you about your comments in part (a)?
```{r}
cat("cor(price, age):", cor(insurance$price, insurance$age)[1],end="\n")
```
Answer: This validates my comments above, there is a correlation, but it is weak. After looking ahead below, this is the strongest correlation of the 3, as shown in the plots above.
```{r}
cat("cor(price, bmi):", cor(insurance$price, insurance$bmi)[1],end="\n")
```
Again, there is a slightly positive relationship between bmi and price, but it is less than the relationship between age and price. 
```{r}
cat("cor(price, children):", cor(insurance$price, insurance$children)[1],end="\n")
```
Finally, there is a very slight positive relationship between  children and price, reinforcing what can be seen in the final plot above. 

c. **4 pts** Create box plots of the response, *price*, and the three qualitative predictors *sex*, *smoker*, and *region*. Based on these box plots, does there appear to be a relationship between these qualitative predictors and the response?

*Hint*: Use the given code to convert the qualitative predictors to factors.

```{r}
#make categorical variables into factors
insurance$sex<-as.factor(insurance$sex) #makes female the baseline level
insurance$smoker<-as.factor(insurance$smoker) #makes no the baseline level
insurance$region<-as.factor(insurance$region) #makes northeast the baseline level
```

```{r}
plot(price~sex,data=insurance,main="price vs. sex",col="gray",pch =16)
```
Answer: Based on the plot above, price does not differ greatly between males and females.No or minimal relationship apparent. 
```{r}
plot(price~smoker,data=insurance,main="price vs. smoker",col="gray",pch =16)
```
Based on the plot above, there does look to be a significant different in price for smokers vs non-smokers. There appears to be a realtionship bewtween price and smoker. 

```{r}
plot(price~region,data=insurance,main="price vs. region",col="gray",pch =16)
```
Region does not appear to have a significant difference in price among the values. No or minimal relationship is apparent.  

d. **3 pts** Based on the analysis above, does it make sense to run a multiple linear regression with all of the predictors?
Answer: I would say yes because there are relationships between each of the variables. They may not be strong but they are worth testing a model on, especially since correlation is not causation and does not fully explain the relationship. 

# Question 2: Fitting the Multiple Linear Regression Model [12 points]

Build a multiple linear regression model, named *model1*, using the response, *price*, and all 6 predictors, and then answer the questions that follow:

a. **6 pts** Report the coefficient of determination (R-squared) for the model and give a concise interpretation of this value.
```{r}
model1<-lm(price ~ .,data=insurance)
model1
```
```{r}
cat("Rˆ2:",summary(model1)$r.squared)
```
R-squared of 75%. Cause of this, we can say that 75% of the variation in the insurance prices can be explained by the other variables (predictors) in the model.


b. **6 pts** Is the model of any use in predicting price? Using $\alpha = 0.05$, provide the following elements of the test of overall regression of the model: null hypothesis $H_0$, alternative hypothesis $H_a$, $F$-statistic or $p$-value, and conclusion.
```{r}
summary(model1)
```
Answer:
Elements are, 
Null Hypothesis - B1 = B2 = B3 = ... = B8 = 0
Alternative Hypothesis - At least one of the slope coefficients is not equal to 0. 
Test Statistic of 500.8
p-value of <2.2e-16

Since the p-value is less than the significance level, we can reject the null hypothesis. There is heavy evidence that at least one of the slope coefficients is not equal to 0. The model is of use in predicting price.

# Question 3: Model Comparison [14 points]

a. **5 pts** Assuming a marginal relationship between *region* and *price*, perform an ANOVA $F$-test on the mean insurance prices among the different regions. Using an $\alpha$-level of 0.05, can we reject the null hypothesis that the means of the regions are equal? Please interpret.
```{r}
anova_mod=aov(price ~ region, insurance)
summary(anova_mod)
```
Answer: Yes, we can reject the null hypothesis that the means of the regions are equal. This is because the p-value of the F-test is less than the alpha level. 

b. **5 pts** Now, build a second multiple linear regression model, called *model2*, using *price* as the response variable, and all variables except *region* as the predictors. Conduct a partial $F$-test comparing *model2* with *model1*. What is the partial-F test p-value? Can we reject the null hypothesis that the regression coefficients for *region* variables are zero at an $\alpha$-level of 0.05?
```{r}
model2<-lm(price ~ .-region,data=insurance)
summary(model2)
```
```{r}
#partial f-test
anova(model2, model1)
```
Partial F-Test p-value is 0.09622, and that is greater than our alpha level. We cannot reject the null hypothesis at a 0.05 alpha level. 

c. **4 pts** What can you conclude from 3a and 3b? Do they provide the exact same results?

Answer: They provide different results. 3a looks at relationship between a variable and the response variable while holding other variables constant. 3b looks at how the variable does when we do not hold other variables constant. 


*Note: Please use model1 for all of the following questions.*

# Question 4: Coefficient Interpretation [7 points]

a. **3 pts** Interpret the estimated coefficient of *sexmale* in the context of the problem. *Make sure female is the baseline level for sex. Mention any assumptions you make about other predictors clearly when stating the interpretation.*
```{r}
model1$coefficients[3]
```
Answer: The price of insurance policies for males are $131.31 less than the price of the policy for females, if other variables are held constant.The reference group is female.


b. **4 pts** If the value of the *bmi* in *model1* is increased by 0.01 and the other predictors are kept constant, what change in the response would be expected?
```{r}
model1$coefficients[4] * 0.01
```
Answer: The 0.01 unit increase in bmi, responds with a $3.391935 increase in price, while other variables are held constant.
# Question 5: Confidence and Prediction Intervals [12 points]

a. **6 pts** Compute 90% and 95% confidence intervals (CIs) for the parameter associated with *age* for *model1*. What observations can you make about the width of these intervals?
```{r}
confint(model1,"age",level =0.9)
```
```{r}
confint(model1,"age",level =0.95)
```
The 90% CI is smaller than that of the 95% CI. The number 0 is not in either of the CIs so that implies that the coefficient for age is statistically significant at both CIs. 

b. **3 pts** Using *model1*, estimate the average price for all insurance policies with the same characteristics as the first data point in the sample. What is the 95% confidence interval? Provide an interpretation of your results.
```{r}
data_5b <-insurance[1,1:6]
predict(model1, data_5b,interval="confidence")
```
The 95% confidence interval is $24,143.98 to $26,443.44. Meaning we're 95% sure that the mean price for all the policies fall within the range previously listed. 

c. **3 pts** Suppose that the *age* value for the first data point is increased to 50, while all other values are kept fixed. Using *model1*, predict the price of an insurance policy with these characteristics. What is the 95% prediction interval? Provide an interpretation of your results.
```{r}
data_5b[1] <- 50
predict(model1, data_5b,interval="prediction")
```
Answer: The 95% prediction interval is $21.313.29 to $45,199.23.The actual prediction is $33,256.26, so it falls in the middle of the prediction interval listed. We are 95% sure that the mean price will fall within the range previously listed. 

