---
title: "HW3 Peer Assessment"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

The fishing industry uses numerous measurements to describe a specific fish.  Our goal is to predict the weight of a fish based on a number of these measurements and determine if any of these measurements are insignificant in determining the weigh of a product.  See below for the description of these measurments.  

## Data Description

The data consists of the following variables:

1. **Weight**: weight of fish in g (numerical)
2. **Species**: species name of fish (categorical)
3. **Body.Height**: height of body of fish in cm (numerical)
4. **Total.Length**: length of fish from mouth to tail in cm (numerical)
5. **Diagonal.Length**: length of diagonal of main body of fish in cm (numerical)
6. **Height**: height of head of fish in cm (numerical)
7. **Width**: width of head of fish in cm (numerical)


## Read the data

```{r}
# Import library you may need
library(car)
# Read the data set
fishfull = read.csv("Fish.csv",header=T, fileEncoding = 'UTF-8-BOM')
row.cnt = nrow(fishfull)
# Split the data into training and testing sets
fishtest = fishfull[(row.cnt-9):row.cnt,]
fish = fishfull[1:(row.cnt-10),]
```

*Please use fish as your data set for the following questions unless otherwise stated.*

# Question 1: Exploratory Data Analysis [8 points]

**(a) Create a box plot comparing the response variable, *Weight*, across the multiple *species*.  Based on this box plot, does there appear to be a relationship between the predictor and the response?**

```{r}
boxplot(Weight~Species, data=fish)

```
Written Answer: 
Based on the box plot, there does appear to be a slight relationship between the predictor and the response. Bream, Pike, and Whitefish all have simlar median values. And Parkki, Perch, Roach, and Smelt all seemingly have similar median values as well.

**(b) Create scatterplots of the response, *Weight*, against each quantitative predictor, namely **Body.Height**, **Total.Length**, **Diagonal.Length**, **Height**, and **Width**.  Describe the general trend of each plot.  Are there any potential outliers?**

```{r}
par (mfrow=c(2,3))
for (i in c (3:7)){
  quant_pred = names(fish[i])
  plot(fish[,i], fish$Weight, xlab = quant_pred, ylab = "Weight")
}

```
Written Answer: 
Body.Height plot: There is a positive relationship/trend with, definitely, one outlier but potentially a few more. The relationship does appear to curve a bit.
Total.Length plot: Very similar to Body.Height vs Weight plot. There is a positive trend that appears to curve out towards the ends of the charts. There is one outlier in the same spot, and potentially more along the trendline.
Diagonal.Length: Same as above. Positive relationship that curves towards the ends. One very visible outlier, others are not as clear but still there.
Height: Slight positive trend, less than the prior 3 charts. There are more outliers present in this chart than the previous ones. Due to how much more spread out this one is, the relationship might actually be linear.
Width: Slightly positive trend, that still looks to curve a bit. There are a few more visibly obvious outliers present.

**(c) Display the correlations between each of the quantitative variables.  Interpret the correlations in the context of the relationships of the predictors to the response and in the context of multicollinearity.**

```{r}
cor(fish[-2])

```
Written Answer:
There seems to be strong positive correlation between the predictors and the response. Due to this, multicollinearity is likely to be an issue in a model that uses all of these predictors. 

**(d) Based on this exploratory analysis, is it reasonable to assume a multiple linear regression model for the relationship between *Weight* and the predictor variables?**
Written Answer: 
Yes, it is reasonable to assume a multiple linear regression model for the relationship between the predictors and the response. As we execute the model, we will be able to further assess if addtional transformations are needed.

# Question 2: Fitting the Multiple Linear Regression Model [8 points]

*Create the full model without transforming the response variable or predicting variables using the fish data set.  Do not use fishtest*

**(a) Build a multiple linear regression model, called model1, using the response and all predictors.  Display the summary table of the model.**

```{r}
model1 = lm(Weight~., data = fish)
summary(model1)
```

**(b) Is the overall regression significant at an $\alpha$ level of 0.01? Explain.**
Written Answer: 
Yes, the overall regression is significant at an alpha level of 0.01. This is because the p-value of the F-statistic; it is less than the alpha level. We can reject the null hypothesis.


**(c) What is the coefficient estimate for *Body.Height*? Interpret this coefficient.**
Written Answer: 
Coefficient estimate for Body.Height is -176.87. So, for every unit increase in Body.Height, there is a 176.87 weight unit decrease, holding all other variables constant. 


**(d) What is the coefficient estimate for the *Species* category Parkki? Interpret this coefficient.**
Written Answer: 
Coefficient estimate for SpeciesParkki is 79.34. So, if a fish is of the species Parkki, the expected weight of the fish is 79.34 units greater than the intercept species, holding all other variables constant.



# Question 3: Checking for Outliers and Multicollinearity [6 points]

**(a) Create a plot for the Cook's Distances. Using a threshold Cook's Distance of 1, identify the row numbers of any outliers.**

```{r}
cook_plot = cooks.distance(model1)
plot(cook_plot) + abline(h = 1)
```
```{r}
cook_plot>1
which(cook_plot>1)
```
Written Answer: 
Row number 30 is an outlier.

**(b) Remove the outlier(s) from the data set and create a new model, called model2, using all predictors with *Weight* as the response.  Display the summary of this model.**

```{r}
fish1 = fish[-30,]
model2 = lm(Weight~., data = fish1)
summary(model2)
```

**(c) Display the VIF of each predictor for model2. Using a VIF threshold of max(10, 1/(1-$R^2$) what conclusions can you draw?**

```{r}
max(10, 1/(1-summary(model2)$r.squared))

```
```{r}
vif(model2)
```
Written Answer:
Based on the VIF threshold, there is high multicollinearity in the model for the predictors.


# Question 4: Checking Model Assumptions [6 points]

*Please use the cleaned data set, which have the outlier(s) removed, and model2 for answering the following questions.*

**(a) Create scatterplots of the standardized residuals of model2 versus each quantitative predictor. Does the linearity assumption appear to hold for all predictors?**

```{r}
resid = rstandard(model2)
par (mfrow=c(2,3))
for (i in c(3:7)){
  quant_pred = names(fish1[i])
  plot(fish1[,i], resid, xlab = quant_pred, ylab = "Standardized Residuals") + abline (h=0)
}
```
Written Answer: 
The linearity assumption does not appear to hold for all predictors. All of the plots show a bit of curvature or non-linearity.


**(b) Create a scatter plot of the standardized residuals of model2 versus the fitted values of model2.  Does the constant variance assumption appear to hold?  Do the errors appear uncorrelated?**

```{r}
plot(model2$fitted.values, resid, xlab = "Fitted", ylab = "Standardized Residuals") + abline(h = 0)
```
Written Answer: 
There appears to be some correlation, in the shape of a V or U. Constant Variance assumption does not hold because the spread of the residuals is not equal across fitted values.


**(c) Create a histogram and normal QQ plot for the standardized residuals. What conclusions can you draw from these plots?**

```{r}
hist(resid)

```
```{r}
qqPlot(resid)
```
Written answer: 
Histogram seems to have a normal distribution but the right tail is longer. The QQ plot shows that right tail that can be seen in the histogram. From these charts, I can assume that the normality assumption does not hold so further transformations may be required.

# Question 5: Partial F Test [6 points]

**(a) Build a third multiple linear regression model using the cleaned data set without the outlier(s), called model3, using only *Species* and *Total.Length* as predicting variables and *Weight* as the response.  Display the summary table of the model3.**

```{r}
model3 = lm(Weight~Species + Total.Length, data = fish1)
summary(model3)
```

**(b) Conduct a partial F-test comparing model3 with model2. What can you conclude using an $\alpha$ level of 0.01?**

```{r}
anova(model3, model2)

```
Written Answer: 
p-value is greater than the alpha level so we cannot reject the null hypothesis.


# Question 6: Reduced Model Residual Analysis and Multicollinearity Test [7 points]

**(a) Conduct a multicollinearity test on model3.  Comment on the multicollinearity in model3.**
```{r}
max(10, 1/(1-summary(model3)$r.squared))

```
```{r}
vif(model3)
```
Written Answer: 
Based on the VIF threshold, there is not much multicollinearity left in the model so we should proceed. 

**(b) Conduct residual analysis for model3 (similar to Q4). Comment on each assumption and whether they hold.**
```{r}
resid2 = rstandard(model3)
par (mfrow=c(2,2))
  plot(fish1$Total.Length, resid2, xlab = quant_pred, ylab = "Standardized Residuals") + abline (h=0) + plot(model3$fitted.values, resid2, xlab = "Fitted", ylab = "Residuals") + abline(h = 0)
```
```{r}
hist(resid2)
```

```{r}
qqPlot(resid2)
```
Written Answer: 
Linearity assumption: Does not hold based on the width vs standardized residuals
Constant Variance assumption: Does not hold based on the residuals vs fitted values
Normality assumption: Does not hold based on the histogram and qqplot

# Question 7: Transformation [9 pts]

**(a) Use model3 to find the optimal lambda, rounded to the nearest 0.5, for a Box-Cox transformation on model3.  What transformation, if any, should be applied according to the lambda value?  Please ensure you use model3**

```{r}
coxbox = boxCox(model3)

```
```{r}
optimal_lambda = coxbox$x[which.max(coxbox$y)]
optimal_lambda
round(optimal_lambda/0.5)*0.05
```
Written Answer: 
Square root transformation of the response should be done.

**(b) Based on the results in (a), create model4 with the appropriate transformation. Display the summary.**
```{r}
model4 = lm(Weight^(1/2)~Species + Total.Length, data = fish1)
summary(model4)
```



**(c) Perform Residual Analysis on model4. Comment on each assumption.  Was the transformation successful/unsuccessful?**
```{r}
resid3 = rstandard(model4)
par (mfrow=c(2,2))
  plot(fish1$Total.Length, resid3, xlab = quant_pred, ylab = "Standardized Residuals") + abline (h=0) + plot(model4$fitted.values, resid3, xlab = "Fitted", ylab = "Residuals") + abline(h = 0)

```
```{r}
hist(resid3)
```
```{r}
qqPlot(resid3)
```
Written Answer:
Linearity assumption: Does hold true. The standardized residuals are better scattered across the 0 line.
Constant Variance assumption: Does hold true. No pattern or trend, as well as no clustering. The U/V shape looks to have gone away.
Normality assumption: QQplot and histogram still show a tail, but the tail is much smaller than before. The normality assumption holds.


# Question 8: Model Comparison [2 pts]

**(a) Using each model summary, compare and discuss the R-squared and Adjusted R-squared of model2, model3, and model4.**
Written Answer: 
R-squared
model2 = 0.9385
model3 = 0.9353
model4 = 0.9817
Model4 has the greatest R-squared value. There a different amount of variables included in these models so comparing R-squared values here does not add value.

Adjusted R-Squared
model2 = 0.9335
model3 = 0.9321
model4 = 0.9808
Model4, again, has the greatest R-squared value. model4 is the best model to use, it explains the most variance. 


# Question 9: Prediction [8 points]

**(a) Predict Weight for the last 10 rows of data (fishtest) using both model3 and model4.  Compare and discuss the mean squared prediction error (MSPE) of both models.** 

```{r}
predm3 = predict(model3,fishtest)
mse.model3 = mean((predm3-fishtest$Weight)^2)
mse.model3
```
```{r}
predm4 = predict(model4,fishtest)^2
mse.model4 = mean((predm4-fishtest$Weight)^2)
mse.model4
```
Written Answer: 
Lower MSPE wins, so model4 is the best model to use. 

**(b) Suppose you have found a Perch fish with a Body.Height of 28 cm, and a Total.Length of 32 cm. Using model4, predict the weight on this fish with a 90% prediction interval.  Provide an interpretation of the prediction interval.**

```{r}
perch_fish = data.frame(Species = "Perch", Total.Length = 32)
predict(model4, perch_fish, interval = "prediction", level = 0.9)^2
```
Written Answer: 
Based on model4, the preferred model from the answers above, the weight of the fish is expected to be 461.9429. The 90% prediction interval is 374.4536 to 558.6091. There is a 90% chance that the weight of the fish will fall in this range.


